---
title: "Forecasting Travellers entering Canada"
author: "Julián Alejandro Úsuga Ortiz"
date: "12/12/2022"
#toc: true
format:
  html:
#    html-math-method: katex
    code-tools: false
    echo: false
#    self-contained: true
execute:
  warning: false
---

# Goal

In this project the goal was trying to predict the number of travellers entering Canada one year into the future. 

Since we only have data up until October 2022, we predict the number of travellers for the next 12 months after this date.

The data was obtained from **Statistics Canada**, [Table 24-10-0050-01 Non-resident visitors entering Canada, by country of residence](https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=2410005001).


# Results

The model is created with all the data and the predicted number of travellers are plotted along with the previous months.

As we can see the COVID-19 pandemic and lockdowns generated a significant drop in the number of travellers.
The predictions show us a potential recovery in the number of travellers and tourism.


```{r}
library(dygraphs)
library(tidyverse)
library(forecast)
load(file = "data/data.RData")
```

```{r}
# data transforming

data.ts <- ts(data$visitors, start = c(1990, 1), frequency = 12)
```

```{r}
# model

len.total <- 1:(length(data.ts) + 12)

tc.ago.2021 <- cbind(I1 = stats::filter(
  as.integer(seq_along(len.total) == 380),
  filter = 0.5,
  method = "rec",
  sides = 1
))


cov.df <- data.frame(
  ls.sep.2001 = as.integer(seq_along(len.total) >= 141),
  ls.mar.2020 = as.integer(seq_along(len.total) >= 363),
  ls.abr.2020 = as.integer(seq_along(len.total) >= 364),
  ao.may.2020 = as.integer(seq_along(len.total) == 365),
  ao.jun.2020 = as.integer(seq_along(len.total) == 366),
  ao.jul.2020 = as.integer(seq_along(len.total) == 367),
  ao.ago.2020 = as.integer(seq_along(len.total) == 368),
  ao.sep.2020 = as.integer(seq_along(len.total) == 369),
  ao.oct.2020 = as.integer(seq_along(len.total) == 370),
  ao.nov.2020 = as.integer(seq_along(len.total) == 371),
  ao.dic.2020 = as.integer(seq_along(len.total) == 372),
  ao.ene.2021 = as.integer(seq_along(len.total) == 373),
  ao.feb.2021 = as.integer(seq_along(len.total) == 374),
  ao.mar.2021 = as.integer(seq_along(len.total) == 375),
  ao.abr.2021 = as.integer(seq_along(len.total) == 376),
  ao.may.2021 = as.integer(seq_along(len.total) == 377),
  ao.jun.2021 = as.integer(seq_along(len.total) == 378),
  ao.jul.2021 = as.integer(seq_along(len.total) == 379),
  tc.ago.2021 = tc.ago.2021
  )
```


```{r}
model.arima <-
  auto.arima(data.ts, seasonal = TRUE, 
             xreg = as.matrix(cov.df[1:length(data.ts),]))

start.pred <- length(data.ts) - 1
end.pred <- start.pred + 12
pred.arima <- forecast(model.arima, xreg =
                          as.matrix(cov.df[start.pred:end.pred, ]))
```


```{r}

# 80 % c.i

data.plot <-
  cbind(data.ts,
        lwr = pred.arima$lower[, 1],
        upr = pred.arima$upper[, 1],
        pred = pred.arima$mean)


dygraph(data.plot, main = "Travellers that visited Canada") %>%
  dySeries(label = "Travellers") %>%
  dyAxis("y", label = "Travellers (in M people)") %>%
  dySeries(c("lwr", "pred", "upr"), label = "Prediction") %>%
  dyRangeSelector(dateWindow = c("2000-01-01", "2023-10-01"))
```


# Predictions

The predictions of the model expressed in millions of travellers. eg. for February of 2023 the expected number of travellers entering Canada is **988787** with a 80% confidence interval of **(733628, 1243947)**.

```{r}
knitr::kable(pred.arima)
```

